---
title: "DDSAnalytics Study"
author: "Kevin Albright"
date: "7/21/2021"
output:
  prettydoc::html_pretty:
    theme: cayman
---

##DDSAnalytics is an analytics company specializing in talent management solutions for Fortune 100 companies. 
 Goal is to conduct analysis of existing employee data to highlight turnover
```{r echo = FALSE, warning=FALSE, message=FALSE}
#install.packages("tidyverse")
#install.packages("ggthemes")
#install.packages("GGally")
#install.packages("viridis")
#install.packages("class")
#install.packages("caret")
#install.packages("prettydoc")
#install.packages("kable_extra")
#install.packages("plotly")
#install.packages("e1071")
#install.packages("randomForest")
library(class)
library(randomForest)
library(scales)
library(caret)
library(e1071)
library(class)
library(plotly)
library(kableExtra)
library(caret)
library(e1071)
library(viridis)
library(ggthemes)
library(GGally)
library(usdata)
library(tidyverse)
library(dplyr)
library(sqldf)
library(prettydoc)
knitr::opts_chunk$set(echo = TRUE)

```
##Load Case Study 2 data set
```{r fig.width=10,fig.height=7,warning=FALSE, message=FALSE}
#Read in Attrition Data
Attrition = read.csv("/Users/Kevin/Desktop/School/Doing Data Science/Project 2/CaseStudy2_data.csv", header = TRUE)
NOAttrition = read.csv("/Users/Kevin/Desktop/School/Doing Data Science/Project 2/CaseStudy2CompSet No Attrition.csv", header = TRUE)
AttritionNo = sqldf('
                    select *
                    from Attrition
                    where "Attrition" = "No"
                    ')
AttritionYes = sqldf('
                    select *
                    from Attrition
                    where "Attrition" = "Yes"
                    ')
```
Parse and Summarize data
```{r fig.width=10,fig.height=7,warning=FALSE, message=FALSE}
Attrition %>%
  group_by(Attrition) %>%
  summarize(MeanCompanysWorked = mean(NumCompaniesWorked),
            MeanMonthlyIncome = mean(MonthlyIncome),
            MeanYearsAtCompany = mean(YearsAtCompany),
            MeanWorkAge = mean(Age),
            MeanTotalWorkingYears = mean(TotalWorkingYears),
            MeanJobLevel = mean(JobLevel),
            MeanPerformance = mean(PerformanceRating),
            MeanYearsSincePromotion = mean(YearsInCurrentRole),
            MeanSalaryHike = mean(PercentSalaryHike),
            Total = n()) %>%
  arrange(Attrition,Total)
```
Plot relationships
```{r fig.width=10,fig.height=7,warning=FALSE, message=FALSE}
AttritionStats = Attrition %>%
  group_by(Attrition) %>%
  summarize(
            MeanAge = mean(Age),
            MeanMonthlyIncome = mean(MonthlyIncome),
            MeanWorkingYears = mean(TotalWorkingYears),
            MeanJobLevel = mean(JobLevel),
            MeanPerformance = mean(PerformanceRating),
            MeanYearsSincePromotion = mean(YearsInCurrentRole),
            MeanSalaryHike = mean(PercentSalaryHike),
            Total = n()) %>%
  arrange(Attrition,Total)
kable(AttritionStats,position = "left")

TravelStats = Attrition %>%
  group_by(Attrition,BusinessTravel) %>%
  summarize(TotalPop= n(),
            MeanWorkAge = mean(Age),
            MeanWorkLifeBalance = mean(WorkLifeBalance),
            MeanJobSatisfaction = mean(JobSatisfaction),
            MeanTotalWorkingYears = mean(TotalWorkingYears),
            MeanMonthlyIncome = mean(MonthlyIncome),
            MeanYearsAtCompany = mean(YearsAtCompany),
            MeanTotalWorkingYears = mean(TotalWorkingYears),
            MeanJobLevel = mean(JobLevel),
            MeanPerformance = mean(PerformanceRating),
            MeanYearsSincePromotion = mean(YearsInCurrentRole),
            MeanSalaryHike = mean(PercentSalaryHike),
            Total = n()) %>%
  arrange(Attrition,Total)
kable(TravelStats,position = "left")



JobRoleStats = Attrition %>%
  group_by(Attrition,JobRole) %>%
  summarize(TotalPop= n(),
            MeanWorkAge = mean(Age),
            MeanWorkLifeBalance = mean(WorkLifeBalance),
            MeanJobSatisfaction = mean(JobSatisfaction),
            MeanTotalWorkingYears = mean(TotalWorkingYears),
            MeanMonthlyIncome = mean(MonthlyIncome),
            MeanYearsAtCompany = mean(YearsAtCompany),
            MeanTotalWorkingYears = mean(TotalWorkingYears),
            MeanJobLevel = mean(JobLevel),
            MeanPerformance = mean(PerformanceRating),
            MeanYearsSincePromotion = mean(YearsInCurrentRole),
            MeanSalaryHike = mean(PercentSalaryHike),
            Total = n()) %>%
  arrange(Attrition,Total)
kable(JobRoleStats,position = "left")

OvertimeStats = Attrition %>%
  group_by(Attrition,OverTime) %>%
  summarize(TotalPop= n(),
            MeanWorkAge = mean(Age),
            MeanTotalWorkingYears = mean(TotalWorkingYears),
            MeanMonthlyIncome = mean(MonthlyIncome),
            MeanYearsAtCompany = mean(YearsAtCompany),
            MeanTotalWorkingYears = mean(TotalWorkingYears),
            MeanJobLevel = mean(JobLevel),
            MeanPerformance = mean(PerformanceRating),
            MeanYearsSincePromotion = mean(YearsInCurrentRole),
            MeanSalaryHike = mean(PercentSalaryHike),
            Total = n()) %>%
  arrange(Attrition,Total)
kable(OvertimeStats,position = "left")


Attrition %>% ggplot(aes(x = Attrition,y = TotalWorkingYears)) + 
  geom_boxplot(color = "blue", fill = "black") + ggtitle("Attrition based on Years in Workforce")

Attrition %>% ggplot(aes(x = Attrition,y = MonthlyRate)) + 
  geom_boxplot(color = "blue", fill = "black") + ggtitle("Attrition based on Monthly Rate") 

Attrition %>% ggplot(aes(x = Attrition,y = YearsInCurrentRole)) + 
  geom_boxplot(color = "blue", fill = "black") + ggtitle("Attrition based on Years in Current Role")

TravelStats %>% ggplot(aes(x = Attrition, y = TotalPop, fill=BusinessTravel)) + 
  geom_bar(stat="identity") + ggtitle("Attrition based on Travel")

TravelStats %>% 
  ggplot(aes(x = reorder(BusinessTravel,TotalPop), y = TotalPop, fill=Attrition)) + 
  geom_bar(stat="identity") + coord_flip() +
  ggtitle("Business Travel Turnover")  + xlab("Travel") + ylab("Total Population")

OvertimeStats %>% ggplot(aes(x = Attrition, y = TotalPop,fill = OverTime)) + geom_bar(stat="identity") + 
  ggtitle("Overtime vs No Overtime") 

Attrition %>% ggplot(aes(x=Age, y=MonthlyIncome)) + 
  geom_point(aes(x=Age, y=MonthlyIncome, color=Attrition))+
  ggtitle('Age vs Monthly Income') +
  scale_color_discrete(name="Attrition") + 
  facet_wrap(~Attrition, scales="free")

Attrition %>% ggplot(aes(x = reorder(JobRole,MonthlyIncome), y = MonthlyIncome,fill=JobRole))+
  geom_boxplot()+
  coord_flip()  +
  ggtitle('Job Role vs Monthly Income') +xlab("Job Role") + ylab("Monthly Income")

JobRoleStatsYes = JobRoleStats %>% filter(Attrition=="Yes") %>% arrange(-TotalPop)
JobRoleStatsNo = JobRoleStats %>% filter(Attrition=="No") %>% arrange(-TotalPop)

JobRoleStats %>% ggplot(aes(x = reorder(JobRole,MeanJobSatisfaction), y = MeanJobSatisfaction, fill=JobRole))+
  geom_boxplot()+
  coord_flip()  +
  ggtitle('Job Role vs Satisfaction') +xlab("Job Role") + ylab("Job Satisfaction")

JobRoleStats %>% ggplot(aes(x = reorder(JobRole,MeanWorkLifeBalance), y = MeanWorkLifeBalance, fill=JobRole))+
  geom_boxplot()+
  coord_flip()  +
  ggtitle('Job Role vs Work Life Balance') +xlab("Job Role") + ylab("Work Life Balance")

JobRoleStatsYes %>% ggplot(aes(x = reorder(JobRole,TotalPop), y = TotalPop, fill=JobRole)) + geom_bar(stat="identity") + 
  coord_flip() + ggtitle("Attrition by Job Role") + xlab("Job Role") + ylab("Total Population")

JobRoleStatsYes %>% ggplot(aes(x = reorder(JobRole,-MeanWorkLifeBalance), y = MeanWorkLifeBalance, fill=JobRole)) + geom_bar(stat="identity") + 
  coord_flip() + ggtitle("Work Life Balance by Job Role") + xlab("Job Role") + ylab("Total Population")

JobRoleStatsNo %>% ggplot(aes(x = reorder(JobRole,TotalPop), y = TotalPop, fill=JobRole)) + geom_bar(stat="identity") + 
  coord_flip() + ggtitle("Attrition by Job Role") + xlab("Job Role") + ylab("Total Population")

JobRoleStatsNo %>% ggplot(aes(x = reorder(JobRole,-MeanWorkLifeBalance), y = MeanWorkLifeBalance, fill=JobRole)) + geom_bar(stat="identity") + 
  coord_flip() + ggtitle("Work Life Balance by Job Role") + xlab("Job Role") + ylab("Total Population")

p <- plot_ly(Attrition, x = ~TotalWorkingYears, y = ~Age, z = ~MonthlyIncome, color = ~JobRole) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'WorkingYears'),
                      yaxis = list(title = 'Age'),
                      zaxis = list(title = 'MonthyIncome')))
p

```
#Perform Classification using KNN
```{r fig.width=10,fig.height=7,warning=FALSE, message=FALSE}

#Model Data using KNN for Age and Monthly Income

  ##Plot Relationship Betwen Age and Monthly Income
Attrition %>% ggplot(aes(x = Age,y=MonthlyIncome,color = Attrition)) + geom_point() + xlab("Age") + geom_smooth(method="lm") + ylab("Monthly Income") + 
  ggtitle("Relationship Between Age and Monthly Income") +
  scale_y_continuous(label=comma)

#Split out training/test data - 70/30
set.seed(100)
splitPerc = .70
trainAttrition= sample(1:dim(Attrition)[1],round(splitPerc * dim(Attrition)[1]))
train = Attrition[trainAttrition,]
test = Attrition[-trainAttrition,]

accs = data.frame(accuracy = numeric(90), k = numeric(90))

#Formulate the optimal k value for KNN
for(i in 1:90)
{
  classifications = knn(train[,c(2,20)],test[,c(2,20)],train$Attrition, prob = TRUE, k = i)
  table(test$Attrition,classifications)
  CM = confusionMatrix(table(test$Attrition,classifications))
  accs$accuracy[i] = CM$overall[1]
  accs$k[i] = i
}
plot(accs$k,accs$accuracy, type = "l", xlab = "k") 
abline(v=accs$k[which.max(accs$accuracy)], col="red")
accs$k[which.max(accs$accuracy)]

splitPerc = .70
trainAttrition= sample(1:dim(Attrition)[1],round(splitPerc * dim(Attrition)[1]))
train = Attrition[trainAttrition,]
test = Attrition[-trainAttrition,]

classification = knn(Attrition[,c(2,20)],Attrition[,c(2,20)],Attrition$Attrition,prob = TRUE, k = 5)

table(classification,Attrition$Attrition)
confusionMatrix(table(classification,Attrition$Attrition))
```
#Perform Classification using Random Forest
```{r fig.width=10,fig.height=7,warning=FALSE, message=FALSE}

#Model Data with Random Forest
  ##Read in Data
Attrition = read.csv("/Users/Kevin/Desktop/School/Doing Data Science/Project 2/CaseStudy2_data.csv", header = TRUE)
NOAttrition= read.csv("/Users/Kevin/Desktop/School/Doing Data Science/Project 2/CaseStudy2CompSet No Attrition.csv", header = TRUE)

#Ensure Attrition is changed to factor
Attrition$Attrition =  factor(as.character(Attrition$Attrition), levels=c("Yes", "No"))

Attrition_Variables = randomForest(Attrition~ .-MonthlyIncome, 
                                   data=Attrition, ntree=1000, 
                                   keep.forest=FALSE,
                                   importance=TRUE)
varImpPlot(Attrition_Variables)



#Split test and train data - 70/30
set.seed(3033)
splitPerc = .70
trainAttrition= sample(1:dim(Attrition)[1],round(splitPerc * dim(Attrition)[1]))
train = Attrition[trainAttrition,]
test = Attrition[-trainAttrition,]

#Apply Random Forest using Monthly Income to test data
EmpAtt = randomForest(Attrition ~ .-Age, 
                      data=train, 
                      strata=train$Attrition, 
                      sampsize= c(60,60))

  #Use newly trained data set to predict test set
AttPredict = predict(EmpAtt, 
                     newdata= test)

  #Create confusion matrix to assess accuracy stats
confusionMatrix(AttPredict, test$Attrition)


#Apply Random Forest to the output file
EmpAtt2 = randomForest(Attrition ~ .-Age, 
                       data=Attrition, 
                       strata=Attrition$Attrition, 
                       sampsize= c(60,60))

AttPredict2 = predict(EmpAtt2, 
                      newdata= NOAttrition)





EmpAttPreds = data.frame(NOAttrition$ID, AttPredict2)
#EmpAttPreds

#write.csv(EmpAttritionPreds, "/Users/Kevin/Desktop/School/Doing Data Science/Project 2/Case2PredictionsAlbrightAttrition.csv")

```
Linear Regression Model vs Random Forest for Predicting Salary
```{r fig.width=10,fig.height=7,warning=FALSE, message=FALSE}
#Linear Regression Model 
NOSalary= read.csv("/Users/Kevin/Desktop/School/Doing Data Science/Project 2/CaseStudy2CompSet No Salary.csv", header = TRUE)

ggplot(data = Attrition, aes(x = Age, y = MonthlyIncome)) +
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  theme(panel.background = element_rect(fill = "white"),
        axis.line.x=element_line(),
        axis.line.y=element_line()) +
  ggtitle("Linear Model Fitted to Data") +
  scale_y_continuous(label=comma)

set.seed(100)
splitPerc = .70
trainAttrition= sample(1:dim(Attrition)[1],round(splitPerc * dim(Attrition)[1]))
trainSalary = Attrition[trainAttrition,]
testSalary = Attrition[-trainAttrition,]


fit_1 <- lm(MonthlyIncome ~ Age, 
            data = trainSalary)

summary(fit_1)

ggplot(data = Attrition, aes(x = Age, y = MonthlyIncome)) +
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  theme(panel.background = element_rect(fill = "white"),
        axis.line.x=element_line(),
        axis.line.y=element_line()) +
  ggtitle("Linear Model Fitted to Data") +
  scale_y_continuous(label=comma)

SalaryPreds = predict(fit_1, NOSalary)

RMSE(Attrition$MonthlyIncome, SalaryPreds)
#data.frame(NOSalary$ID,SalaryPreds)


#Random Forest Model
  #Train Data

splitPerc = .70
trainAttrition= sample(1:dim(Attrition)[1],round(splitPerc * dim(Attrition)[1]))
trainSalary = Attrition[trainAttrition,]
testSalary = Attrition[-trainAttrition,]
str(trainSalary)

EmpRandoSalary = randomForest(MonthlyIncome ~ .-Age, 
                              data=trainSalary)

PredictSalary = predict(EmpRandoSalary, 
                          newdata= testSalary)
  #Test Data
RMSE(testSalary$MonthlyIncome, PredictSalary)
PredDF = data.frame(MonthlyIncome=testSalary$MonthlyIncome, PredictSalary)
PredDF %>% ggplot(aes(x=MonthlyIncome,y=PredictSalary)) + geom_point(aes(x=MonthlyIncome,y=PredictSalary)) + 
  ggtitle("Relationship Between Predicted and Actuals") + xlab("Actual") + ylab("Predicted") + geom_smooth(method="lm") 
#data.frame(testSalary$ID, PredictSalary)

#RMSE of Random Forest model for predicting salary
NOSalary$Attrition =  factor(as.character(NOSalary$Attrition), levels=c("Yes", "No"))
SalaryRF <- randomForest(MonthlyIncome ~ .-Age, data=trainSalary)
#importance(SalaryRF)
#varImpPlot(SalaryRF)

#Test on provided salary data set
#str(NOSalary)
#str(Attrition)
PredictSalary2<- predict(SalaryRF, 
                         newdata= NOSalary,
                         importance=TRUE,ntree=500)
summary(PredictSalary2)

PredictionsAlbright <- data.frame(NOSalary$ID, PredictSalary2)
write.csv(PredictionsAlbright,"/Users/Kevin/Desktop/School/Doing Data Science/Project 2/AlbrightSalaryPredictions.csv")

